Rotation of Magnetic Islands in Tokamak Plasmas

### 2022-01-27 Research/NonlinearTearingModeTheory/NonlinearTearingModeTheory.tex
Latest version

\section{Introduction}
Tearing modes are slowly growing instabilities of ideally-stable tokamak plasmas that reconnect magnetic field-lines
at various resonant surfaces within the plasma, in the process forming magnetic island chains that degrade the plasma confinement.\cite{wes} {\em Mode locking}\/ is a process by which the rotation of a magnetic island chain is braked  due to electromagnetic interaction with the resistive 
vacuum vessel surrounding the plasma, causing the chain to eventually {\em lock}\/ (i.e., become stationary in the laboratory frame) to an error-field.\cite{nave}
 Locked magnetic island chains are one of the principal causes of disruptions in tokamaks.\cite{vries}

It is well known that single-fluid, resistive magnetohydrodynamics (MHD) offers a very poor description of
tearing mode dynamics in tokamak plasmas. 
For instance, the strong {\em diamagnetic}\/ flows present in such plasmas decouple the electron and ion flows to some extent,
necessitating a two-fluid treatment.\cite{ara} Previously, Fitzpatrick \& Waelbroeck\,\cite{fw,fw1} used a generalized version of the four-field model of Hazeltine, et al.\cite{haz})
to determine the two-fluid response of a nonlinear magnetic island chain to a resistive vacuum vessel  in a large aspect-ratio tokamak plasma. The aim of this paper is to revisit this analysis in the light of certain refinements to magnetic island theory
introduced in Ref.~\onlinecite{rf2021}.

@@

As is well known, the general solution of Eq.~(\ref{e100}) is such that $\delta\psi$ is continuous across the resonant surface,
whereas $\partial\delta\psi/\partial r$ is discontinuous. The discontinuity in $\partial \delta\psi/\partial r$ implies the presence of a helical  current sheet  at the resonant surface. 
The complex quantity 
\begin{equation}\label{e108}
{\mit\Psi}_s(t)= \delta\psi(r_s,t) 
\end{equation}
parameterizes the amplitude and phase of the
reconnected magnetic flux at the resonant surface.\cite{rf1993}
The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \delta\psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface.

@@

Simultaneously matching our separable solution of Eq.~(\ref{e3})  across the resonant surface and the wall yields\,\cite{rf1993}
\begin{align}
{\mit\Delta\Psi}_s& = E_{ss}\,{\mit\Psi}_s + E_{sw}\,{\mit\Psi}_w,\\[0.5ex]
{\mit\Delta\Psi}_w& = E_{ws}\,{\mit\Psi}_s+ E_{ww}\,{\mit\Psi}_w.
\end{align}
Here, $E_{ss}= [r\,d\hat{\psi}_s/dr]_{r_{s-}}^{r_{s+}}$,
$E_{ww}= [r\,d\hat{\psi}_w/dr]_{r_{w-}}^{r_{w+}}$,
$E_{sw} =[r\,{d\hat\psi}_w/dr]_{r=r_{s+}}$, and 
$E_{ws} =-[r\,{d\hat\psi}_s/dr]_{r=r_{w-}}$
are real quantities determined by the solutions of Eqs.~(\ref{e100}) and (\ref{e100a}).
It is easily demonstrated that 
$E_{sw} = E_{ws}$.\cite{rf1993}

Standard electromagnetic theory applied to the resistive wall reveals that\,\cite{nave,rf1993}
\begin{equation}\label{e9}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
$\tau_w= \mu_0\,r_w\,\delta_w/\eta_w$.
Here, $\eta_w$ and $\delta_w$ are the electrical resistivity and radial thickness  of the wall, respectively. 
In deriving Eq.~(\ref{e9}), we have made use of the so-called {\em thin wall}\/ approximation, according to
which the radial variation of $\delta\psi$ across the wall is negligible.  The thin wall approximation is
valid as long as
\begin{equation}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

Assuming an $\exp(-{\rm i}\,\omega\,t)$ time dependence of perturbed quantities, the previous four equations
can be combined to give
\begin{equation}\label{e11}
\frac{{\mit\Delta\Psi}_s}{{\mit\Psi}_s}= {\mit\Delta}_{pw} + \frac{{\mit\Delta}_{nw}-{\mit\Delta}_{pw}}{1-{\rm i}\,\omega\,\tau_{LR}},
\end{equation}
where ${\mit\Delta}_{pw}= E_{ss}$ is the  tearing stability index\cite{fkr} when the wall is perfectly conducting,
${\mit\Delta}_{nw} = E_{ss} + E_{sw}^{\,2}/(-E_{ww})$ is the  tearing stability index when there is no wall,
and $\tau_{LR}= \tau_w/(-E_{ww})$ is the effective L/R time of the wall. (Note that $E_{ww}<0$.)

\section{Resonant Plasma Response Model}\label{sfour}
\subsection{Introduction}
The helical current sheet at the resonant surface can only be resolved by solving  a two-fluid, resistive-MHD, plasma response model
in the inner region (i.e., the region of the plasma in the immediate vicinity of the resonant surface), 
and asymptotically matching the solution so obtained to the ideal-MHD solution in the outer region (i.e., everywhere else in the plasma). The particular 
plasma response model used in this paper is
described in this section.

@@

\subsection{Resonant Response Model}\label{resonant}
The fundamental fields in our resonant plasma reponse model---namely, $\psi$, $\phi$, $N$, $V$, and $J$---have the following
definitions:
\begin{align}\label{e10}
\nabla\psi &= \frac{{\bf n}\times{\bf B}} {r_s\,B_z},\\[0.5ex]
\nabla\phi &= \frac{{\bf n}\times {\bf V}}{r_s\,V_A},\\[0.5ex]
N &=-\hat{d}_i\left(\frac{p-p_0}{B_z^{\,2}/\mu_0}\right),\\[0.5ex]
V &= \hat{d}_i\left(\frac{{\bf n}\cdot {\bf V}}{V_A}\right),\label{e13}\\[0.5ex]
J&=-\frac{2\,\epsilon_s}{q_s}+\hat{\nabla}^2\psi.
\end{align}
Here,   ${\bf n} = (0,\,\epsilon/q_s,\,1)$, $\epsilon = r/R_0$, $q_s=m/n$, 
$V_A =B_z/\sqrt{\mu_0\,n_0\,m_i}$, 
$d_i = \sqrt{m_i/(n_0\,e^{\,2}\,\mu_0)}$,  $\hat{d}_i=d_i/r_s$, 
$\epsilon_s=r_s/R_0$, and $\hat{\nabla} = r_s\,\nabla$.

Our resonant plasma response model takes the form:\,\cite{fw}
\begin{align}
\frac{\partial\psi}{\partial\hat{t}}&= [\phi,\psi] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\psi]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial \hat{\nabla}^2\phi}{\partial \hat{t}}&= [\phi,\hat{\nabla}^2\phi] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\psi] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right), \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\psi]
+ \hat{D}_\parallel \,[[N,\hat{\psi}],\hat{\psi}]+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\psi] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V.\label{e21}
\end{align}
Here, $[A,B]\equiv \hat{\nabla} A\times \hat{\nabla} B\cdot {\bf n}$, $\hat{t} = t/(r_s/V_A)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,r_s\,V_A)$, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, 
$\hat{\chi}_\varphi= \chi_\varphi/(r_s\,V_A)$, where $\eta_{\parallel,\perp}$ is the parallel/perpendicular plasma electrical
resistivity at the resonant surface, $E_\parallel$ the parallel inductive electric field that maintains the equilibrium toroidal
plasma current in the vicinity of the resonant surface, and $\chi_\varphi$  the (anomalous) perpendicular ion momentum
diffusivity at the resonant surface. 
Moreover, $d_\beta=c_\beta\,d_i$, and $\hat{d}_\beta=d_\beta/r_s$, where $c_\beta = \sqrt{\beta/(1+\beta)}$, and
$\beta=(5/3)\,\mu_0\,p_0/B_z^{\,2}$. Here, $d_\beta$ is usually referred to as the {\em ion sound radius}. Furthermore,
$\lambda_e=0.71\,\eta_e/(1+\eta_e)$. Finally,\cite{fw}
\begin{align}
\hat{D}_\parallel&\equiv\frac{D_{\parallel}}{r_s\,V_A}= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp& \equiv \frac{D_{\perp}}{r_s\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right),
\end{align}
where $\hat{\chi}_{\parallel\,e,i}=\chi_{\parallel\,e,i}/(r_s\,V_A)$,  $\hat{\chi}_{\perp\,e,i}=\chi_{\perp\,e,i}/(r_s\,V_A)$, 
$\chi_{\parallel\,e,i}$ is the parallel electron/ion energy diffusivity in the vicinity of the resonant surface, and 
$\chi_{\perp\,e,i}$ is the anomalous
electron/ion perpendicular energy diffusivity in the vicinity of the resonant surface.

@@

It is convenient  to work in a frame of reference that corotates with the tearing mode. In this reference frame, the
reconnected flux, ${\mit\Psi}_s$, is assumed to be a positive real
quantity. If $W$ is the full radial width of the magnetic island chain that develops at the resonant surface then it is
helpful to define the reduced island width:
\begin{equation}\label{e26a}
w = \frac{W}{4}= \left(\frac{L_s\,{\mit\Psi}_s}{B_z}\right)^{1/2}.
\end{equation}
It is assumed that $w\ll r_s$. Let $\hat{w}=w/r_s$. 
 In the limit $|\hat{x}|/\hat{w}\gg 1$ (i.e., many island widths from the resonant surface), we have  
\begin{align}\label{e26}
\psi &\rightarrow \frac{\hat{x}^{\,2}}{2\,\hat{L}_s} + \frac{{\mit\Psi}_s}{r_s\,B_z}\,\cos\zeta,\\[0.5ex]
\phi &\rightarrow \left(\frac{\hat{\omega}}{m}- \hat{V}_E\right)\hat{x} - \frac{s}{4}\,\hat{V}_E'\,\hat{x}^{\,2},\\[0.5ex]
N &\rightarrow -\hat{V}_\ast\,\hat{x},\\[0.5ex]
V&\rightarrow \hat{V}_\parallel,\\[0.5ex]
J& \rightarrow -\frac{2}{q_s\,\hat{R}_0} + \frac{1}{\hat{L}_s}.\label{e30}
\end{align}
where 
$\hat{x}=(r-r_s)/r_s$,
 $\hat{L}_s=L_s/r_s$,  $L_s=R_0\,q_s/s_s$, $\hat{R}_0=R_0/r_s$, $\zeta=m\,\theta-n\,\varphi$, 
  $\hat{V}_E= V_E(r_s)/V_A$,
$\hat{V}_\ast= V_\ast(r_s)/V_A$,
$V_\ast(r) = (dp/dr)/(e\,n_0\,B_z)$ 
is the (dominant $\theta$-component of the) diamagnetic velocity,
 $\hat{V}_\parallel=\hat{d}_i\, V_z(r_s)/V_A$, $s_s=s(r_s)$, and $s(r)=d\ln q/d\ln r$. 
Here, $\omega = \hat{\omega}\,V_A/r_s$ is the rotation frequency of the tearing mode  in the
laboratory frame,
$s = {\rm sgn}(\hat{x})$,
and 
\begin{equation}
\hat{V}_E' = \frac{1}{V_A}\left[r\,\frac{dV_E}{dr}\right]_{r_{s-}}^{r_{s+}}.
\end{equation}
The parameter $\hat{V}_E'$ is introduced  into the analysis in order to take into account the fact that the plasma rotation profile in the outer region develops a
gradient discontinuity at the resonant surface in response to the nonlinear localized electromagnetic torque that emerges 
at the surface.\cite{fw} Finally, 
 $\hat{E}_\parallel =(2/s_s-1) \,(\hat{\eta}_\parallel/\hat{L}_s)$.

>>>>

\section{Introduction}
Tearing modes are slowly growing instabilities of ideally-stable tokamak plasmas that reconnect magnetic field-lines
at various resonant surfaces within the plasma leading to the formation of magnetic island chains that degrade the plasma confinement.\cite{wes} {\em Mode locking}\/ is a process by which the rotation of a magnetic island chain is braked  due to electromagnetic interaction with the resistive 
vacuum vessel surrounding the plasma, causing the chain to eventually {\em lock}\/ (i.e., become stationary in the laboratory frame) to an error-field.\cite{nave}
 Locked magnetic island chains are one of the principal causes of disruptions in tokamaks.\cite{vries}

It is well known that single-fluid, resistive magnetohydrodynamics (MHD) offers a very poor description of
tearing mode dynamics in tokamak plasmas. 
For instance, the strong {\em diamagnetic}\/ flows present in such plasmas decouple the electron and ion flows,
necessitating a two-fluid treatment.\cite{ara} Previously, Fitzpatrick \& Waelbroeck\,\cite{fw,fw1} used a generalized version of the four-field model of Hazeltine, et al.\cite{haz})
to determine the two-fluid response of a nonlinear magnetic island chain to a resistive vacuum vessel  in a large aspect-ratio tokamak plasma. The aim of this paper is to revisit this analysis in the light of certain refinements to magnetic island theory
introduced in Ref.~\onlinecite{rf2021}.

@@

As is well known, the general solution of Eq.~(\ref{e100}) is such that $\delta\psi$ is continuous across the resonant surface,
whereas $\partial\delta\psi/\partial r$ is discontinuous.\cite{fkr} The discontinuity in $\partial \delta\psi/\partial r$ implies the presence of a helical  current sheet  at the resonant surface. 
The complex quantity 
\begin{equation}\label{e108}
{\mit\Psi}_s(t)= \delta\psi(r_s,t) 
\end{equation}
parameterizes the amplitude and phase of the
reconnected magnetic flux at the resonant surface.\cite{rf1993}
The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \delta\psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface.

@@

Simultaneously matching our separable solution of Eq.~(\ref{e3})  across the resonant surface and the wall yields\,\cite{rf1993}
\begin{align}\label{e9a}
{\mit\Delta\Psi}_s& = E_{ss}\,{\mit\Psi}_s + E_{sw}\,{\mit\Psi}_w,\\[0.5ex]
{\mit\Delta\Psi}_w& = E_{ws}\,{\mit\Psi}_s+ E_{ww}\,{\mit\Psi}_w.
\end{align}
Here, $E_{ss}= [r\,d\hat{\psi}_s/dr]_{r_{s-}}^{r_{s+}}$,
$E_{ww}= [r\,d\hat{\psi}_w/dr]_{r_{w-}}^{r_{w+}}$,
$E_{sw} =[r\,{d\hat\psi}_w/dr]_{r=r_{s+}}$, and 
$E_{ws} =-[r\,{d\hat\psi}_s/dr]_{r=r_{w-}}$
are real quantities determined by the solutions of Eqs.~(\ref{e100}) and (\ref{e100a}).
It is easily demonstrated that 
$E_{sw} = E_{ws}$.\cite{rf1993}

Standard electromagnetic theory applied to the resistive wall reveals that\,\cite{nave,rf1993}
\begin{equation}\label{e9}
{\mit\Delta\Psi}_w = \tau_w\,\frac{d{\mit\Psi}_w}{dt},
\end{equation}
where
$\tau_w= \mu_0\,r_w\,\delta_w/\eta_w$ is known as the {\em time constant}\/ of the wall.
Here, $\eta_w$ and $\delta_w$ are the electrical resistivity and radial thickness  of the wall, respectively. 
In deriving Eq.~(\ref{e9}), we have made use of the so-called {\em thin wall}\/ approximation, according to
which the radial variation of $\delta\psi$ across the wall is negligible.  The thin wall approximation is
valid as long as
\begin{equation}\label{e10}
\frac{r_w}{\delta_w}\gg \tau_w\left|\frac{d\ln{\mit\Psi}_w}{dt}\right|.
\end{equation}

Assuming an $\exp(-{\rm i}\,\omega\,t)$ time dependence of perturbed quantities,  Eqs.~(\ref{e9a})--(\ref{e9}) 
can be combined to give
\begin{equation}\label{e11}
\frac{{\mit\Delta\Psi}_s}{{\mit\Psi}_s}= {\mit\Delta}_{pw} + \frac{{\mit\Delta}_{nw}-{\mit\Delta}_{pw}}{1-{\rm i}\,\tau_{LR}\,\omega},
\end{equation}
where ${\mit\Delta}_{pw}= E_{ss}$ is the  so-called {\em tearing stability index}\,\cite{fkr} when the wall is perfectly conducting,
${\mit\Delta}_{nw} = E_{ss} + E_{sw}^{\,2}/(-E_{ww})$ is the  tearing stability index when there is no wall,
and $\tau_{LR}= \tau_w/(-E_{ww})$ is the effective L/R time of the wall. (Note that $E_{ww}<0$.)

\section{Resonant Plasma Response Model}\label{sfour}
\subsection{Introduction}
In order to proceed further, we need a relationship between ${\mit\Delta\Psi}_s$ and ${\mit\Psi}_s$ that
is analogous in form to Eq.~(\ref{e9}). Such a relationship specifies how the helical current sheet that
flows at the resonant surface is related to the reconnected magnetic flux. 
We can obtain the desired relationship  by solving  a two-fluid, resistive-MHD, plasma response model
in the inner region (i.e., the region of the plasma in the immediate vicinity of the resonant surface), 
and asymptotically matching the solution so obtained to the ideal-MHD solution in the outer region (i.e., everywhere else in the plasma). The particular 
plasma response model used in this paper is
described in this section.

@@

\subsection{Resonant Response Model}\label{resonant}
The fundamental (dimensionless) fields in our resonant plasma reponse model---namely, $\psi$, $\phi$, $N$, $V$, and $J$---have the following
definitions:
\begin{align}\label{e10}
\nabla\psi &= \frac{{\bf n}\times{\bf B}} {r_s\,B_z},\\[0.5ex]
\nabla\phi &= \frac{{\bf n}\times {\bf V}}{r_s\,V_A},\\[0.5ex]
N &=-\hat{d}_i\left(\frac{p-p_0}{B_z^{\,2}/\mu_0}\right),\\[0.5ex]
V &= \hat{d}_i\left(\frac{{\bf n}\cdot {\bf V}}{V_A}\right),\label{e13}\\[0.5ex]
J&=-\frac{2\,\epsilon_s}{q_s}+\hat{\nabla}^2\psi.
\end{align}
Here,   ${\bf n} = (0,\,\epsilon/q_s,\,1)$, $\epsilon = r/R_0$, $q_s=m/n$, 
$V_A =B_z/\sqrt{\mu_0\,n_0\,m_i}$, 
$d_i = \sqrt{m_i/(n_0\,e^{\,2}\,\mu_0)}$,  $\hat{d}_i=d_i/r_s$, 
$\epsilon_s=r_s/R_0$, and $\hat{\nabla} = r_s\,\nabla$.

Our resonant plasma response model takes the form:\,\cite{fw}
\begin{align}
\frac{\partial\psi}{\partial\hat{t}}&= [\phi,\psi] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\psi]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial \hat{\nabla}^2\phi}{\partial \hat{t}}&= [\phi,\hat{\nabla}^2\phi] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\psi] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right), \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\psi]
+ \hat{D}_\parallel \,[[N,\hat{\psi}],\hat{\psi}]+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\psi] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V.\label{e21}
\end{align}
Here, $[A,B]\equiv \hat{\nabla} A\times \hat{\nabla} B\cdot {\bf n}$, $\hat{t} = t/(r_s/V_A)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,r_s\,V_A)$, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, 
$\hat{\chi}_\varphi= \chi_\varphi/(r_s\,V_A)$, where $\eta_{\parallel,\perp}$ is the parallel/perpendicular plasma electrical
resistivity at the resonant surface, $E_\parallel$ the parallel inductive electric field that maintains the equilibrium toroidal
plasma current in the vicinity of the resonant surface, and $\chi_\varphi$  the (anomalous) perpendicular ion momentum
diffusivity at the resonant surface. 
Moreover, $d_\beta=c_\beta\,d_i$, and $\hat{d}_\beta=d_\beta/r_s$, where $c_\beta = \sqrt{\beta/(1+\beta)}$, and
$\beta=(5/3)\,\mu_0\,p_0/B_z^{\,2}$. Here, $d_\beta$ is usually referred to as the {\em ion sound radius}. Furthermore,
$\lambda_e=0.71\,\eta_e/(1+\eta_e)$. Finally,\cite{fw}
\begin{align}
\hat{D}_\parallel&\equiv\frac{D_{\parallel}}{r_s\,V_A}= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp& \equiv \frac{D_{\perp}}{r_s\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right),
\end{align}
where $\hat{\chi}_{\parallel\,e,i}=\chi_{\parallel\,e,i}/(r_s\,V_A)$,  $\hat{\chi}_{\perp\,e,i}=\chi_{\perp\,e,i}/(r_s\,V_A)$, 
$\chi_{\parallel\,e,i}$ is the parallel electron/ion energy diffusivity in the vicinity of the resonant surface, and 
$\chi_{\perp\,e,i}$ is the (anomalous) perpendicular
electron/ion  energy diffusivity in the vicinity of the resonant surface.

@@

It is convenient  to work in a frame of reference that corotates with the tearing mode. In this reference frame, the
reconnected flux, ${\mit\Psi}_s$, is assumed to be a positive real
quantity. Let $W$ be the full radial width of the magnetic island chain that develops at the resonant surface. It
is
helpful to define the reduced island width:
\begin{equation}\label{e26a}
w \equiv \frac{W}{4}= \left(\frac{L_s\,{\mit\Psi}_s}{B_z}\right)^{1/2}.
\end{equation}
It is assumed that $w\ll r_s$. Let $\hat{w}=w/r_s$. 
 In the limit $|\hat{x}|/\hat{w}\gg 1$ (i.e., many island widths from the resonant surface), we have  
\begin{align}\label{e26}
\psi &\rightarrow \frac{\hat{x}^{\,2}}{2\,\hat{L}_s} + \frac{{\mit\Psi}_s}{r_s\,B_z}\,\cos\zeta,\\[0.5ex]
\phi &\rightarrow \left(\frac{\hat{\omega}}{m}- \hat{V}_E\right)\hat{x} - \frac{\varsigma}{4}\,\hat{V}_E'\,\hat{x}^{\,2},\\[0.5ex]
N &\rightarrow -\hat{V}_\ast\,\hat{x},\\[0.5ex]
V&\rightarrow \hat{V}_\parallel,\\[0.5ex]
J& \rightarrow -\frac{2}{q_s\,\hat{R}_0} + \frac{1}{\hat{L}_s},\label{e30}
\end{align}
where 
$\hat{x}=(r-r_s)/r_s$,
 $\hat{L}_s=L_s/r_s$,  $L_s=R_0\,q_s/s_s$, $\hat{R}_0=R_0/r_s$, $\zeta=m\,\theta-n\,\varphi$, 
  $\hat{V}_E= V_E(r_s)/V_A$,
$\hat{V}_\ast= V_\ast(r_s)/V_A$,
$V_\ast(r) = (dp/dr)/(e\,n_0\,B_z)$ 
is the (dominant $\theta$-component of the) diamagnetic velocity,
 $\hat{V}_\parallel=\hat{d}_i\, V_z(r_s)/V_A$, $s_s=s(r_s)$, and $s(r)=d\ln q/d\ln r$. 
Here, $\omega = \hat{\omega}\,V_A/r_s$ is the rotation frequency of the tearing mode  in the
laboratory frame,
$\varsigma = {\rm sgn}(\hat{x})$,
and 
\begin{equation}
\hat{V}_E' = \frac{1}{V_A}\left[r\,\frac{dV_E}{dr}\right]_{r_{s-}}^{r_{s+}}.
\end{equation}
The parameter $\hat{V}_E'$ is introduced  into our analysis in order to take into account the fact that the plasma rotation profile in the outer region develops a
gradient discontinuity at the resonant surface in response to the nonlinear localized electromagnetic torque that emerges 
at the surface due to interaction with the resistive wall.\cite{fw} Finally, 
 $\hat{E}_\parallel =(2/s_s-1) \,(\hat{\eta}_\parallel/\hat{L}_s)$.

\section{Nonlinear Solution of Resonant Plasma Response Model}
\subsection{Rescaled Response Model}\label{s8.2}
Let
$X = \hat{x}/\hat{w}$,
and
$T = \omega_\ast\,t$,  where $\omega_\ast = - (m/r_s)\,V_\ast(r_s)$. 
It follows that $|X|\sim {\cal O}(1)$ in the immediate vicinity of the island chain. It is helpful to define the
rescaled fields ${\mit\Psi}(X,\zeta,T)$, ${\mit\Phi}(X,\zeta,T)$, ${\cal N}(X,\zeta,T)$,
${\cal V}(X,\zeta,T)$, and ${\cal J}(X,\zeta,T)$, where
\begin{align}
\psi&= \left(\frac{\hat{w}^{\,2}}{\hat{L}_s}\right){\mit\Psi},\label{e466}\\[0.5ex]
\phi &=\left(\frac{\hat{\omega}_\ast\,\hat{w}}{m}\right){\mit\Phi},\\[0.5ex]
N &= \left(\frac{\hat{\omega}_\ast\,\hat{w}}{m}\right){\cal N},\\[0.5ex]
V&= \hat{V}_\parallel + \left(\frac{\hat{L}_s\,\hat{\omega}_\ast^{\,2}}{m^2\,c_\beta^{\,2}}\right){\cal V},\\[0.5ex]
J & =-\frac{2}{q_s\, \hat{R}_0} + \frac{1}{\hat{L}_s} + \left(\frac{\hat{L}_s\,\hat{\omega}_\ast^{\,2}}{m^2\,\hat{w}^{\,2}}\right){\cal J},
\end{align}
and  $\hat{\omega}_\ast = \omega_\ast/(V_A/r_s)$. 
The resonant response model specified in Sect.~\ref{resonant} rescales to give
\begin{align}\label{e469}
\frac{d(\ln \hat{w}^2)}{dT}\,\cos\zeta &= \left\{{\mit\Phi}-\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,{\cal N}, {\mit\Psi}\right\}
+ \epsilon_\beta\,\epsilon_d\,\epsilon_\eta\,{\cal J},\\[0.5ex]
\frac{\partial(\partial_X^{\,2}\,{\mit\Phi})}{\partial T} &= \partial_X\!\!\left\{{\mit\Phi} + \frac{{\cal N}}{1+\tau},\partial_X{\mit\Phi}\right\}
+ \{{\cal J},{\mit\Psi}\}+ \epsilon_\chi \,\partial_X^{\,4}\!\left({\mit\Phi} + \frac{{\cal N}}{1+\tau}\right),\label{e470}\\[0.5ex]
\frac{\partial {\cal N}}{\partial T} &= \{{\mit\Phi}, {\cal N}\} + \{{\cal V},{\mit\Psi}\} + \epsilon_d\,\epsilon_n\,\{{\cal J},{\mit\Psi}\}\nonumber\\[0.5ex]&\phantom{=}
+ \epsilon_D \!\left[\left(\frac{w}{w_c}\right)^4\,\{\{{\cal N},{\mit\Psi}\}, {\mit\Psi}\}+\partial_X^{\,2}{\cal N}\right],\label{e471}\\[0.5ex]
\epsilon_d\,\frac{\partial {\cal V}}{\partial T} &= \epsilon_d\,\{{\mit\Phi},{\cal V}\}+\{{\cal N}, {\mit\Psi}\} + \epsilon_d\,\epsilon_\chi\,\partial_X^{\,2} {\cal V},\label{e472}\\[0.5ex]
\partial_X^{\,2}{\mit\Psi} &= 1+ \epsilon_\beta\,\epsilon_d\,{\cal J}.\label{e473}
\end{align}
Here,  $\partial_X\equiv \partial/\partial X$, 
\begin{equation}
\{A,B\} \equiv \frac{\partial A}{\partial X}\,\frac{\partial B}{\partial\zeta} - \frac{\partial A}{\partial \zeta}\,\frac{\partial B}{\partial X},
\end{equation}
and
\begin{equation}
w_c = \left(\frac{1}{\epsilon_s\,s_s\,n}\right)^{1/2}\left(\frac{D_\perp}{D_\parallel}\right)^{1/4}r_s.
\end{equation}
Furthermore,
\begin{align}
\epsilon_d &= \left(\frac{L_s}{L_n}\right)^2\left(\frac{d_\beta}{w}\right)^2,\label{e479}\\[0.5ex]
\epsilon_\beta &= c_\beta^{\,2},\label{e480}\\[0.5ex]
\epsilon_n&= \left(\frac{L_n}{L_s}\right)^2,\\[0.5ex]
\epsilon_\eta &= \frac{\eta_\parallel}{\mu_0\,\omega_\ast\,w^2},\label{e481}\\[0.5ex]
\epsilon_\chi &= \frac{\chi_\phi}{\omega_\ast\,w^2},\label{e482}\\[0.5ex]
\epsilon_D &= \frac{D_\perp}{\omega_\ast\,w^2}.
\end{align}
Finally,
\begin{equation}
L_n= \left[-\frac{1}{{\mit\Gamma}\,(1-c_\beta^{\,2})}\left(\frac{d\ln n}{dr}\right)_{r_s}\right]^{-1}
\end{equation}
is the {\em density gradient scalelength}\/  at the resonant surface.

Equations~(\ref{e469})--(\ref{e473}) must be solved subject to the boundary conditions [cf.\ Eqs.~(\ref{e26})--(\ref{e30})] 
\begin{align}
{\mit\Psi}&\rightarrow \frac{X^{2}}{2} + \cos\zeta,\label{e483}\\[0.5ex]
{\mit\Phi} &\rightarrow v\,X +\frac{\varsigma\,v'\,X^{2}}{2},\label{e484}\\[0.5ex]
{\cal N}&\rightarrow X,\label{e486}\\[0.5ex]
{\cal V}&\rightarrow 0,\label{e486a}\\[0.5ex]
{\cal J}& \rightarrow 0,\label{e487}
\end{align}
as $|X|\rightarrow\infty$.  Here,
\begin{align}
v &= \frac{\omega-\omega_E}{\omega_\ast},\label{e488}\\[0.5ex]
v' &= -\frac{\hat{w}}{2\,\omega_\ast}\left[r\,\frac{d\omega_E}{dr}\right]_{r_{s-}}^{r_{s+}},\label{e489}
\end{align}
where $\omega_E= (m/r_s)\,V_E(r_s)$. 
Note that ${\mit\Psi}$, ${\mit\Phi}$, ${\cal N}$, ${\cal V}$, and ${\cal J}$ are all ${\cal O}(1)$ quantities.
Note, further,  that the boundary conditions (\ref{e483})--(\ref{e487}),
as well as the symmetry of Eqs.~(\ref{e469})--(\ref{e473}), ensure that ${\mit\Psi}$, ${\cal V}$, and
${\cal J}$ are even functions of $X$, whereas ${\mit\Phi}$ and ${\cal N}$ are odd functions.
